---
title: "Variable Work"
author: "Devin Shackle"
date: "11/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install/Load Libraries

```{r}
#Check to make sure that all the libraries we need are installed
packages <- c("data.table"
              , "dplyr"
              , "GGally"
              , "ggplot2"
              , "Hmisc"
              , "knitr"
              , "reshape2"
#              , "RODBC"
              , "scales"
              , "stringr")
for (package in packages) {
  message(sprintf("Checking that %s is installed...", package))
  if (!package %in% row.names(installed.packages())) {
    install.packages(package, repos = "https://cran.rstudio.com")
  }
}

# Now load all the libraries we need into our environment
library(data.table)
library(dplyr)
library(GGally)
library(ggplot2)
library(Hmisc)
library(knitr)
library(reshape2)
#library(RODBC)
library(scales)
library(stringr)


```


## Let's Get the Data
```{r}

## Modify this path to point to the data set you would like to use
## Get the 100K data from Dropbox
PATH_TO_DATA = '~/Dropbox/MSAn Capstone - Fall 2017/Data/3. 100K Sampled Data/FINAL2_SAMPLED100.csv'

## Form the total data frame
df <- read.csv(PATH_TO_DATA)

## Remove objects that are no longer needed
rm(package)
rm(packages)
rm(PATH_TO_DATA)

## Split the data into two peices; this will keep on the match data
df.m1 <- df[df$MATCH==1,]
```

## Data Cleaning: Remove variables with too much missing data
If a variable has too many missing values or if it is almost entirely made up of one value, let's get rid of it
```{r}


## Check the data in the column
#  Remove the variables from the data if 90% of the data in the column is the same value
eor <- nrow(df.m1)[1]
eoc <- dim(df.m1)[2] - 1
drop.vars = c()

## Compute the mode of a factor vector
Mode <- function(v){
  uniqv <- na.omit(unique(v))
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

## Remove the variables where 90% of the values are common or are missing
#  Will use the mode function to determine the most common value
#  Will begin at the ninth variable as there are the start of the third party data
#  The last variable is 'MATCH' variable and should not be included
for(i in 9:eoc){
  
  var.mode <- Mode(df.m1[,i])
  len.na <- sum(is.na(df.m1[,i]))
  len.mode <- length(df.m1[df.m1[,i]==var.mode,i])
  if(len.na > len.mode){
    len.num <- len.na
  } else {
    len.num <- len.mode
  }
  
  if(len.num/eor >= .90){
    if (length(drop.vars) == 0){
      drop.vars <-c(names(df.m1[i]))  
    } else {
      drop.vars <-c(drop.vars, names(df.m1[i]))
    }
    
  }
}

## Reveal dropped variables for examination
drop.vars

#  Evaluate other variables that having little value
bypassVars <- c(
    "V4"      # <- Does not match with State
  , "V38"     # <- Can't see how this variable helps
  , "Zip4"    # <- Can't see how this variable helps
)

#  Combine the two lists
drop.vars <- Reduce(union, list(drop.vars, bypassVars))


## Reduce df.m1 for the variables that failed the 90% test
# Ref: https://stackoverflow.com/questions/4605206/drop-data-frame-columns-by-name
df.m1.r <- df.m1[ , !(names(df.m1) %in% drop.vars)]


## Remove objects that are no longer needed
rm(eoc)
rm(eor)
rm(i)
rm(len.mode)
rm(len.na)
rm(len.num)
rm(var.mode)
rm(bypassVars)
rm(drop.vars)
rm(Mode)
#rm(df.m1)
```


## PART I
Classify the variables to a specific data type
```{r}
## This part of the code is soley intended to treat variables having responses that require a
#  treatment before use.  These include variables in which "blank" means something.  These variables
#  will be set to 99 (as a category).  There are also variables having responses that are "yes" or
#  "no" that will be converted to 1's or 0's.  The outcome of this section will channel directly
#  into a missing variable treatment operation.

isBlnkVar <- c(
    "Upmarket"
  , "V35"
  , "V42"
  , "V43"
  , "V44"
  , "V45"
  , "V46"
  , "V47"
  , "V48"
  , "V49"
  , "V50"
  , "V51"
  , "V52"
  , "V53"
  , "V54"
  , "V55"
  , "V56"
  , "V57"
  , "V65"
  , "V74"
  , "V80"
  , "V81"
  , "V85"
  , "V86"
  , "V160"
)

## A note about the response variables
#  There are no missing values in RESP
#  It does not require any special handing

isRespVar <- c(
  "RESP"
)

#  Zeros in these variables will be converted to nulls; zeros are meaningless
isZeroVar <- c(
    "V14" 
  , "V69"
  , "V81"
  , "V82"
  , "V88"
  , "V90"
  , "V92"
  , "V93"
  , "V134"
  , "V136"
  , "V140"
)

#  Yes in these variables will be converted to 1
isYesNoVar <- c(
    "V71"
  , "V119"
  , "V163"
  , "V164"
  , "V165"
  , "V166"
  , "V167"
  , "V168"
  , "V169"
  , "V170"
  , "V171"
  , "V172"
  , "V173"
  , "V174"
  , "V175"
  , "V176"
  , "V177"
  , "V178"
  , "V179"
  , "V180"
  , "V181"
  , "V182"
  , "V183"
  , "V184"
  , "V185"
  , "V186"
  , "V187"
  , "V188"
  , "V189"
  , "V190"
  , "V191"
  , "V192"
  , "V193"
  , "V194"
  , "V195"
  , "V196"
  , "V197"
  , "V198"
  , "V199"
  , "V200"
  , "V201"
  , "V202"
  , "V203"
  , "V204"
  , "V205"
  , "V206"
  , "V207"
  , "V208"
  , "V209"
  , "V210"
  , "V211"
  , "V212"
  , "V213"
  , "V214"
  , "V215"
  , "V216"
  , "V217"
  , "V218"
  , "V219"
  , "V220"
  , "V221"
  , "V222"
  , "V223"
  , "V224"
  , "V225"
  , "V226"
  , "V227"
  , "V228"
  , "V229"
  , "V230"
  , "V231"
  , "V232"
  , "V233"
  , "V234"
  , "V235"
  , "V236"
  , "V237"
  , "V238"
  , "V239"
  , "V240"
  , "V241"
  , "V242"
  , "V243"
  , "V244"
  , "V245"
  , "V246"
  , "V247"
  , "V248"
  , "V249"
  , "V250"
  , "V251"
  , "V252"
  , "V253"
  , "V254"
  , "V255"
  , "V256"
  , "V257"
  , "V258"
  , "V259"
  , "V260"
  , "V261"
  , "V262"
  , "V263"
  , "V264"
  , "V265"
)

#  These variables had leading zeros requiring treatment; they need to be replaced
isPadVar <- c(
  "Zip"
  , "Zip4"
  , "V7"
  , "V12"
  , "V13"
  , "V62"
  , "V68"
  , "V69"
  , "V79"
  , "V80"
  , "V82"
  , "V83"
  , "V285"
  , "V286"
  , "V287"
  , "V288"
)

#  These variables are continuous numeric variables
isNumVar <- c(
  "days_since_last_activity"
  , "V73"
  , "V99"
  , "V101"
  , "V103"
  , "V105"
  , "V106"
  , "V107"
  , "V110"
  , "V113"
  , "V131"
  , "V132"
  , "V272"
  , "V274"
  , "V276"
  , "V278"
)

## Set up a sink file for variable review
sink(file = "df.m1.Part1.txt", append = FALSE, type = c("output", "message"), split = FALSE)

## Some data isn't really missing, the blanks have meaning
# Based upon the reading, these variables need a conversion to character from factor to work
# The first six variables should not have blanks

# Decided to use a loop as the function seemed complex (for now)
# The lesson learned here is the when refering to a column by i, use i
# https://stackoverflow.com/questions/3969852/update-data-frame-via-function-doesnt-work
```


```{r}
unique(df.m1$MATCH)

#  One 176 of the original data variables remain at this point, but this sequence will add
#  additional dummy variables as required
#  Remove variables categories from the previous stage.
rm(df.m1.r)
rm(isBlnkVar)
rm(isNumVar)
rm(isPadVar)
rm(isYesNoVar)
rm(isZeroVar)
#rm(package)
#rm(packages)

## The following is a list of the variables the team decided would be dropped.
#  This list was provided by Sandra on 10/28/2017 15:36
isDropVar <- c(
  "V2"
, "V7"
, "V9"
, "V11"
, "V12"
, "V13"
, "V14"
, "V29"
, "V30"
, "V36"
, "V37"
, "V39"
, "V41"
, "V58"
, "V60"
, "V62"
, "V65"
, "V70"
, "V71"
, "V75"
, "V76"
, "V77"
, "V78"
, "V79"
, "V81"
, "V82"
, "V83"
, "V84"
, "V87"
, "V89"
, "V91"
, "V92"
, "V93"
, "V94"
, "V96"
, "V97"
, "V100"
, "V102"
, "V104"
, "V114"
, "V115"
, "V118"
, "V120"
, "V133"
, "V135"
, "V137"
, "V139"
, "V141"
, "V142"
, "V143"
, "V145"
, "V146"
, "V147"
, "V149"
, "V151"
, "V153"
, "V162"
, "V269"
, "V285"
, "V286"
, "V287"
, "V288"
)

#  This list is added to account for problem variables
#  They contain mulitple values (i.e. GT 100) that I can't see are needed
isProbVar <- c(
  "V80"
, "V279"
, "Zip"
, "Zip4"
)

```


## Prepare the funtions

```{r}
#  Create a function to discect the year from date variables
datFun <- function(X){
  as.integer(substr(X, 1, 4))
}

#  Create a function to process histograms
histFun <- function(df, X){
  
  # par(mfrow=c(1,2))
  # hist(df[,2][df[,1]==0], main=NULL, xlab="RESP = 0", col=gray.colors(12))
  # hist(df[,2][df[,1]==1], main=NULL, xlab="RESP = 1", col=gray.colors(12))
  # title(main=paste("Histogram of",X), outer=TRUE, line=-2)
  # par()
  
  par(mfrow=c(1,2))
  hist(df[,2][df[,2]>min(df[,2])], main=NULL, xlab="BEFORE", col=gray.colors(12))
  hist(df[,3][df[,3]>min(df[,2])], main=NULL, xlab="AFTER", col=gray.colors(12))
  title(main=paste("Histogram of",X), outer=TRUE, line=-2)
  par()
}
```


## PART III
Reclassify the variables to a specific data type
```{r}
#  This part of the code is intended to reclassify variables having into variables types that will be
#  used during modeling.  The variable assignments are to those that were determined during the campus
#  immersion.

## Define character variables
isCharVar <- c(
  "State"
, "V15"
, "V80"
, "V85"
, "V86"
, "V109"
, "V111"
, "V116"
, "V117"
, "V159"
, "V267"
, "V268"
, "V270"
, "V271"
, "V273"
, "V275"
, "V277"
, "V279"
)

## Introduce a new variable classification object for dates
isDateVar <- c(
  "V32"
, "V61"
, "V98"
, "V108"
, "V112"
)

## Define factor variables
isFacVar <- c(
  "Census_Block"
, "V35"
, "V40"
, "V43"
, "V44"
, "V45"
, "V46"
, "V47"
, "V59"
, "V68"
, "V69"
, "V74"
, "V88"
, "V90"
, "V134"
, "V136"
, "V138"
, "V140"
, "V148"
, "V284"
, "V289"
, "V290"
, "V291"
, "V292"
, "V293"
, "V294"
, "V295"
, "V296"
)

## Define integer variables - these have 0 or 1 values
isIntVar <- c(
  "Upmarket"
, "Match"
, "V95"
, "V119"
, "V152"
, "V163"
, "V164"
, "V165"
, "V166"
, "V167"
, "V168"
, "V169"
, "V170"
, "V171"
, "V172"
, "V173"
, "V174"
, "V175"
, "V176"
, "V177"
, "V178"
, "V179"
, "V180"
, "V181"
, "V182"
, "V183"
, "V184"
, "V185"
, "V186"
, "V187"
, "V188"
, "V189"
, "V190"
, "V191"
, "V192"
, "V193"
, "V194"
, "V195"
, "V196"
, "V197"
, "V198"
, "V199"
, "V200"
, "V201"
, "V202"
, "V203"
, "V204"
, "V205"
, "V206"
, "V207"
, "V208"
, "V209"
, "V210"
, "V211"
, "V212"
, "V213"
, "V214"
, "V215"
, "V216"
, "V217"
, "V218"
, "V219"
, "V220"
, "V221"
, "V222"
, "V223"
, "V224"
, "V225"
, "V226"
, "V227"
, "V228"
, "V229"
, "V230"
, "V231"
, "V232"
, "V233"
, "V234"
, "V235"
, "V236"
, "V237"
, "V238"
, "V239"
, "V240"
, "V241"
, "V242"
, "V243"
, "V244"
, "V245"
, "V246"
, "V247"
, "V248"
, "V249"
, "V250"
, "V251"
, "V252"
, "V253"
, "V254"
, "V255"
, "V256"
, "V257"
, "V258"
, "V259"
, "V260"
, "V261"
, "V262"
, "V263"
, "V264"
, "V265"
)

## Define numeric variables
isNumVar <- c(
  "days_since_last_activity"
, "V73"
, "V99"
, "V101"
, "V103"
, "V105"
, "V106"
, "V107"
, "V110"
, "V113"
, "V121"
, "V122"
, "V123"
, "V124"
, "V126"
, "V127"
, "V128"
, "V129"
, "V130"
, "V131"
, "V132"
, "V144"
, "V150"
, "V272"
, "V274"
, "V276"
, "V278"
)

## Provide for the expected additional variables
isPlusVar <- c(
  "V32.d"
, "V61.d"
, "V98.d"
, "V108.d"
, "V112.d"
)

isLogVar <- c(
  "V99"
, "V101"
, "V103"
, "V105"
, "V106"
, "V107"
, "V113"
, "V131"
, "V132"
)

## It's assumed that the following variables are kept following the review of the individual
#  variables against the response variable.  The following is a complete variable list of 
#  those the team decided to keep.  It is constructed from the individual lists above.
df.vars.keep <- Reduce(union, list(
    isRespVar
  , isCharVar
  , isDateVar
  , isFacVar
  , isIntVar
  , isNumVar
  )
)

## Ensure there are no variables in the 'keep' variables that are considered dropped
#  df.vars.chk <- df.m1[, (names(df.m1) %in% isDropVar)]
#  names(df.vars.chk)

## Rebuild 'df.m1' by removing the drop variables provided by Sandra
#  Thus, the 'df.m1' data reduces to 141 variables
#  These are in addition to the variables dropped from the previous stage
df.m1.r <- df.m1[, (!names(df.m1) %in% isDropVar & !names(df.m1) %in% isProbVar)]
names(df.m1.r)

## Set up a sink file for variable review
sink(file = "df.m1.Part3.txt", append = FALSE, type = c("output", "message"), split = FALSE)

#  The current class of each variables needs to be checked and modified, if necessary
#  This process will start at the third variable as the first two do not require a conversion
#  - 1: Response; 2: individual_id
#  At the completion of this loop, there should be several new dummy variables, one for each
#  factor and character variable.  Another reduction will occur that removes the original variables.

eor <- dim(df.m1.r)[1]
eoc <- dim(df.m1.r)[2]

for(i in 3:eoc){
  
  if(!names(df.m1.r[i]) %in% isRespVar){
    
    print(paste(names(df.m1.r[i]),"->",class(df.m1.r[,i])))
    print(paste("Variable",i))
    
    if(names(df.m1.r[i]) %in% isDateVar) {
      
      ## Use year function to compute date difference by year
      if(names(df.m1.r[i])=="V32"){
        df.m1.r$V32.d <- (2017 - datFun(df.m1.r$V32))
      }
      if(names(df.m1.r[i])=="V61"){
        df.m1.r$V61.d <- (2017 - datFun(df.m1.r$V61))
      }
      if(names(df.m1.r[i])=="V98"){
        df.m1.r$V98.d <- (2017 - datFun(df.m1.r$V98))
      }
      if(names(df.m1.r[i])=="V108"){
        df.m1.r$V108.d <- (2017 - datFun(df.m1.r$V108))
      }
      if(names(df.m1.r[i])=="V112"){
        df.m1.r$V112.d <- (2017 - datFun(df.m1.r$V112))
      }
    }
    
    if(names(df.m1.r[i]) %in% isCharVar | names(df.m1.r[i]) %in% isFacVar) {
      
        uniques <- unique(df.m1.r[i])
        for (j in uniques) {
          var_name <- paste0(names(df.m1.r[i]),"_Level_",j)
          df.m1.r[,var_name] <- if_else(df.m1.r[i] == i, 1, 0)
        }
    }
    
    if(names(df.m1.r[i]) %in% isNumVar) {
      
      df.m1.r[,i] <- as.numeric(df.m1.r[,i])
    }
    
    if(names(df.m1.r[i]) %in% isIntVar | names(df.m1.r[i]) %in% isDateVar) {
      
      df.m1.r[,i] <- as.integer(df.m1.r[,i])
    }
    
    print(paste(names(df.m1.r[i]),"->",class(df.m1.r[,i])))
    cat("\n")
  }
}

## Process the log transformations
#  They were not processed above as the added variables would not be included in the range
#  of processig variables.  At this point character and factor variables have been converted to
#  indictor variables and will be removed in the following section so that a numeric matrix can
#  be created for the correlation matrix in this later stage.

# eor <- dim(df.m1.r)[1]
# eoc <- dim(df.m1.r)[2]
# 
# for(i in 3:eoc){
#   
#   if(!names(df.m1.r[i]) %in% isRespVar){
#     
#     if(names(df.m1.r[i]) %in% isLogVar | names(df.m1.r[i]) %in% isPlusVar) {
#       
#       eov <- dim(df.m1.r)[2]+1
#       var_name <- paste0("Log_of_",names(df.m1.r[i]))
#       
#       print(summary(df.m1.r)[,i])
#       df.m1.r[var_name] <- df.m1.r[i]
#       df.m1.r[eov] <- log(df.m1.r[eov]+1)
#       print(summary(df.m1.r)[,eov])
#       cat("\n")
#       
#       # Prepare a plot showing the transformation
#       me.df <- data.frame(df.m1.r$RESP, df.m1.r[[i]], df.m1.r[[var_name]])
#       histFun(me.df, var_name)
#     }
#   }
# }

## Remove the original factor and character variables
df.m1.r2 <- df.m1.r[, (!names(df.m1.r) %in% isCharVar & !names(df.m1.r) %in% isFacVar)]
print(names(df.m1.r2))
cat("\n")

## Close sink
sink()

#  Remove variables categories from the previous stage.
rm(df.vars.keep)
rm(eoc)
rm(eor)
rm(i)
rm(j)
rm(var_name)
rm(uniques)
rm(datFun)

## Send the data to a file
#  This data file should include only numeric data where MATCH=1
write.csv(df.m1.r2,'df.m1_Part3.csv', quote = FALSE)
```

## PART IV
This part of the code will prepare a correlation matrix on which to evaluate variable against the response variable

```{r}
#  Force 'df.m1.r2' to a numeric matrix
df.m1.r2 <- mutate_each(df.m1.r2, funs(as.numeric))
names(df.m1.r2)

#  Computes an ordered list of all variables by correlation with response
#  This will also be used for output format and reporting
col_names <- c("Variable","Corr","P.Value","abs.Corr")
var_List <- names(df.m1.r2)[ - which(names(df.m1.r2) == "RESP")]
list <- data.frame(matrix(nrow = length(var_List), ncol = length(col_names)))
names(list) <- col_names

#  Loop through the variables and prepare a corrletion coefficient for each against 'RESP'
#  Prepare the first number for the list sequence
n <- 1

#  This process will start at the third variable as the first two are not requires
#  - 1: Response; 2: individual_id
for (i in 3:(length(var_List))) {
  
  p <- cor.test(df.m1.r2[[var_List[[i]]]], df.m1.r2$RESP)
  list[n,"Variable"] <- var_List[[i]]
  list[n,"Corr"] <- p$estimate
  list[n,"P.Value"] <- p$p.value
  list[n,"abs.Corr"] <- abs(p$estimate)
  n <- n + 1
}

## Set up a sink file for variable review
sink(file = "df.m1.Part4.txt", append = FALSE, type = c("output", "message"), split = FALSE)

#  Construct the output table for evaluation
list <- list[order(-list$abs.Corr),]
kable(list, caption = "Pearson Correlation Coefficients w/RESP")

## Close sink
sink()

#  Remove variables categories from the previous stage.
rm(list)
rm(col_names)
rm(i)
rm(n)
rm(p)
rm(var_List)

## Set up a sink file for warnings
sink(file = "df.m1.warn.txt", append = FALSE, type = c("output", "message"), split = FALSE)

print(warnings())

## Close sink
sink()
```
